\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{titlesec}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{listings}
\usepackage{color}
\usepackage{indentfirst}
\usepackage[normalem]{ulem}

\captionsetup{compatibility=false}
\definecolor{orange}{RGB}{255,72,0}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}

\lstset{frame=tb,
  language=Python,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\color{blue},
  keywordstyle=\color{orange},
  commentstyle=\color{dkgreen},
  stringstyle=\color{dkgreen},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3
}

\hypersetup{
pdftitle={thesis},
pdfauthor={Ciprian},
pdfkeywords={pdf, latex, tex, ps2pdf, pdflatex},
bookmarksnumbered,
pdfstartview={FitH},
urlcolor=cyan,
colorlinks=true,
linkcolor=red,
citecolor=green,
}


\font\myfont=cmr12 at 24pt
\font\myfontE=cmr12 at 18pt
\font\myfontS=cmr12 at 16pt
\title{\textbf{BABEŞ-BOLYAI UNIVERSITY CLUJ-NAPOCA \\ \vspace{1ex} 
FACULTY OF MATHEMATICS AND COMPUTER SCIENCE \\ \vspace{1ex}
SPECIALIZATION COMPUTER SCIENCE \vspace{5ex}}}
\author{\myfont Diploma thesis}
\date{\myfontE Focus detector with Convolutional neural network}

\setcounter{secnumdepth}{4}
\usepackage{natbib}
\usepackage{graphicx}
\titleformat{\paragraph}
{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
\titlespacing*{\paragraph}
{0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}



\begin{document}

\clearpage\maketitle
\thispagestyle{empty}

\vspace{32ex}
{\myfontE {\hspace*{3em}Supervisor:} {\hspace*{8em}Author:}}
\newline
{\myfontE Prof. Dr. DIOSAN Laura {\hspace*{3em}Cuibus Ciprian}}
\begin{center}
    \vspace{3ex}
    {\myfontE 2019}
\end{center}

\pagenumbering{roman}

\textcolor{green}{\sout{Sa adaugi in partea de jos a primei pagini si anul (2019)} si sa incerci sa formatezi prima pagina ca in exemplul de coperta de pe site-ul cs}

\newpage
\begin{center}
    \textit{\textbf{ Abstract}} 
\end{center}
\input{Chapters/Abstract.tex}

\textcolor{green}{\sout{tb si un abstract (o poveste plastica, ne-tehnica, "pt bunica") care sa contina ce problema rezolvi si utilitatea ei, cum o rezolvi, cat de bine ai reusit sa o rezolvi si structura lucrarii (similar cu finalul introducerii, dar poate mai scurt)}}

\textcolor{green}{\sout{cred ca ai putea reorganiza sectiunile astfel:\\
1. Introducere\\
2. AI in automotive (actuala 2.3)\\
3. Theory (actuala 2.1 si 2.2) \\
4. Focus detector (Actuala 2.4, separat in 2: state of art si abordarea ta)\\
5. App (si aici inainte de sectiunea 3.4 Experiments sa ai o sectiune de Testing and validation (asserts, unittesting, coverage, etc.)\\
6. Conclusions}}

\tableofcontents
\listoffigures
\listoftables

\newpage
\pagenumbering{arabic}
\section{Introduction}
    \subsection{Importance \&\ Motivation}
        \input{Chapters/Introduction/MotivationImportance.tex}
    \subsection{Objectives}
        \input{Chapters/Introduction/Objectives.tex}
    \subsection{Thesis structure and original contribution(s)}
        \input{Chapters/Introduction/ThesisStructure.tex}

\newpage
\section{Artificial intelligence in automotive}
    \input{Chapters/TheoreticalPart/Automotive.tex}
\newpage
\section{Theoretical Part}

This chapter will present some historical thoughts about AI, followed by a short overview of machine learning. Moreover, a deep dive is made on neural network, convolutional neural network and how they work.

\textcolor{green}{\sout{arata mai bine daca intre titlul capitolului si prima sectiune ai adauga 2-3 propozitii despre ce e vorba in capitolul curent (un mic rezumat)}}

    \subsection{Introduction to AI}
        \input{Chapters/TheoreticalPart/IntroductionToAI.tex}
    \subsection{Neural network}
        \subsubsection{Background}
            \input{Chapters/TheoreticalPart/NeuralNetwork.tex}
        \subsubsection{Convolutional neural network}            \input{Chapters/TheoreticalPart/ConvolutionalNeuralNerwork.tex}
        
        \subsubsection{CNN Architecture}
        \input{Chapters/TheoreticalPart/CNNArchitecture.tex}
    
        \paragraph{Convolutional layer}
        \input{Chapters/TheoreticalPart/Layers/ConvolutionLayer.tex}
        \paragraph{Pooling layer}
            \input{Chapters/TheoreticalPart/Layers/PoolingLayer.tex}
        \paragraph{Fully-connected layer}
            \input{Chapters/TheoreticalPart/Layers/Fully-ConnectedLayer.tex}
        \paragraph{Backpropagation}
            \input{Chapters/TheoreticalPart/Layers/Backpropagation.tex}
    \subsection{Drowsiness detector}
\newpage
\section{Focus detector}
    \input{Chapters/TheoreticalPart/FocusDetector.tex}
\newpage
\section{App development}
    \subsection{Problem statement + requirements}
        \input{Chapters/PracticalPart/ProblemStatementReq.tex}
    \subsection{Problem design and analysis}
        \subsubsection{Architecture}
           \input{Chapters/PracticalPart/Architecture.tex}
        \subsubsection{Frameworks and setup}
           \input{Chapters/PracticalPart/FrameworksAndSetup.tex}
    \subsection{Implementation}
        \input{Chapters/PracticalPart/Implementation.tex}
    \subsection{Testing}
    \subsection{Experiments}

\newpage
\section{Conclusion}
\bibliographystyle{abbrv}
\begin{thebibliography}{1}
    \bibitem{trafficConflict} 
    Vittorio Astarita* and Vincenzo Pasquale Giofré. 
    \textit{From traffic conflict simulation to traffic crash simulation:  introducing traffic safety indicators based on  the explicit simulation of potential driver errors }. 
    Simulation Modelling Practice and Theory. Issue 94, p:215-236, March 2019.
     
    \bibitem{briefAIHistory} 
    Albert Bruce G. Buchanan. 
    \textit{A (Very) Brief History of Artificial Intelligence}.
    Ai Magazine, 26(4):53-60, December 2005.
    
    \bibitem{MLIntro} 
    Alex Smola and S.V.N. Vishwanathan. 
    \textit{Introduction to Machine Learning}.Cambridge University Press, United Kingdom, 2008.
    
    \bibitem{IntelliAgents}
    Russell, Stuart J.; Norvig, Peter, 
    \textit{Artificial Intelligence: A Modern Approach (2nd ed.)}, Upper Saddle River, New Jersey: Prentice Hal, 2003.
    
    \bibitem{ANNBasic}
    Su-Hyun Han, Ko Woon Kim, SangYun Kim, Young Chul Youn.
    \textit{Artificial Neural Network:
    Understanding the Basic Concepts
    without Mathematics}. Dement Neurocognitive Disord. 2018 Sep;17(3):83-89
    
    \bibitem{clinical}
    Ebersole JS, Husain AM, Nordli DR.
    \textit{Current Practice of Clinical Electroencephalography}. 4th ed. Philadelphia,
    PA: Wolters Kluwer, 2014.
    
    \bibitem{NNandML}
    Haykin S, Haykin SS.
    \textit{Neural Networks and Learning Machines}.
     3rd ed. Upper Saddle River, NJ: Prentice Hall, 2009.
     
    \bibitem{ANNbriefOverview}
    Zakaria M, AL-Shebany M, Sarhan S.
    \textit{Artificial neural network: a brief overview}. Int J Eng Res April 2014;4:7-12.
     
    \bibitem{MakeyourOwn}
    Rashid T. \textit{Make Your Own Neural Network}.
    1st ed. North Charleston, SC: CreateSpace Independent Publishing Platform, 2016.
    
    \bibitem{RealTime}
    Rashid T, Huang BQ, Kechadi MT.
    \textit{ A new simple recurrent network with real-time recurrent learning process}. Proceedings of the 14th Irish Conference on Artificial Intelligence \&\ Cognitive Science, AICS
    2003; 2003 September 17–19; Dublin, Ireland. [place unknown]: Artificial Intelligence and Cognitive Science, 2003;169-174.
    
    \bibitem{ClassifNN}
    Kozyrev SV 
    \textit{ Classification by ensembles of neural networks}. p-Adic Numbers Ultrametric Anal Appl
    2012;4:27-33. 
    
    \bibitem{OnlineApprox}
    Bottou L. Chapter 2. On-line learning and stochastic approximations. In: Saad D, editor. 
    \textit{ On-Line Learning in Neural Networks}.  Cambridge: Cambridge University Press, 1999;9-42.
    
    \bibitem{NNTricks}
    Bottou L. Stochastic gradient descent tricks. In: Montavon G, Orr GB, Müller KR, editors.
    \textit{Neural Networks: Tricks of the Trade}.
    2nd ed. Heidelberg: Springer-Verlag Berlin Heidelberg, 2012;421-436.
     
    \bibitem{IntroCNN}
    K.  O’Shea  and  R.  Nash.
    \textit{An introduction to convolutional neural networks}. 2015.
    
    \bibitem{Largescale}
    Karpathy A., Toderici G., Shetty S., Leung T., Sukthankar R., Fei-Fei L.
    \textit{Largescale video classification with convolutional neural networks}. In: Computer Vision and Pattern Recognition. (CVPR), 2014 IEEE Conference on. pp. 1725–1732. IEEE, 2014.
    
    \bibitem{3dconvolutional}
    Ji S., Xu W., Yang M., Yu K.
    \textit{: 3d convolutional neural networks for human action recognition}.
    Pattern Analysis and Machine Intelligence, IEEE Transactions on 35(1), 221–231, 2013.
    
    \bibitem{EvalPooling}
    D. Scherer, A. Muller, and S. Behnke.
    \textit{Evaluation of pooling operations in convolutional architectures for object recognition}. In International Conference on Artificial Neural Networks, 2010.
    
    \bibitem{AutoAI}
    Keshav Bimbraw. 
    \textit{Autonomous cars: Past, present and future a review of the developments in the last century, the present scenario and the expected future of autonomous vehicle technology}. In Proc. 12th Int. Conf. Inform. Control, Automat. Robot.(ICINCO), vol. 1, Jul 2015, pp.191-198.
    
    \bibitem{Vision}
     A. Geiger, P. Lenz, C. Stiller, and R. Urtasun. 
     \textit{Vision meets robotics:
     The kitti dataset}. International Journal of Robotics Research(IJRR), 2013.
     
    \bibitem{Dlib}
    \href{http://dlib.net/}{Dlib Documentation}. \textit{dlib.net}. Retrieved 2019-05-05.
     
     \bibitem{MIT}
     Lex Fridman
     , Daniel E. Brown, Michael Glazer, William Angell, Spencer Dodd, Benedikt Jenik, Jack Terwilliger, Julia Kindelsberger, Li Ding, Sean Seaman, Hillary Abraham, Alea Mehler, Andrew Sipperley, Anthony Pettinato, Bobbie Seppelt, Linda Angell, Bruce Mehler, Bryan Reimer. 
     \textit{MIT Autonomous Vehicle Technology Study: Large-Scale Deep Learning Based Analysis of Driver Behavior and Interaction with Automation}. arXiv preprint arXiv:1711.06976, 30 Sept 2018.
     
    \bibitem{Numpy}
    \href{https://www.numpy.org/}{Numpy Documentation}. \textit{numpy.org}. Retrieved 2019-05-05.
    
    \bibitem{Frust}
    I. Abdic, L. Fridman, D. McDuff, E. Marchi, B. Reimer, and B. Schuller.
    \textit{Driver frustration detection from audio and video in the wild},in KI 2016: Advances in Artificial Intelligence: 39th Annual German Conference on AI, Klagenfurt, Austria, September 26-30, 2016, Proceedings, vol. 9904. Springer, 2016, p. 237.
    
    \bibitem{DetectEmotional}
    H. Gao, A. Yuce, and J.-P. Thiran.
    \textit{“Detecting emotional stress from facial expressions for driving safety}, in Image Processing (ICIP), 2014 IEEE International Conference on. IEEE, 2014, pp. 5961–5965.
    
    \bibitem{EndTOEnd}
    M. Bojarski, D. Del Testa, D. Dworakowski, B. Firner, B. Flepp, P. Goyal, L. D. Jackel, M. Monfort, U. Muller, J. Zhang et al, 
    \textit{End to end learning for self-driving cars}, arXiv preprint arXiv:1604.07316, 2016.
    
    \bibitem{Tensorflow}
    M. Abadi, A. Agarwal et al.,
    \textit{Tensorflow: Large-scale machine learning on heterogeneous distributed systems,} arXiv:1603.04467, 2016.
    
    \bibitem{KerasDoc}
    \href{https://keras.io/#why-this-name-keras}{Keras Documentation}. \textit{keras.io}. Retrieved 2019-05-05.
    
    \bibitem{KerasBack}
    \href{https://keras.io/backend/}{Keras Backend}. \textit{keras.io}. Retrieved 2019-05-05.
    
    \bibitem{KerasWhy}
    \href{https://keras.io/why-use-keras/}{Keras Why}. \textit{keras.io}. Retrieved 2019-05-05.
    
    \bibitem{OpenCV}
    Kari Pulli, Anatoly Baksheev, Kirill Kornyakov, Victor Eruhimov. \textit{Real-time computer vision with OpenCV}. Commun ACM 2019;55(6):61-9.
    
    \bibitem{SciPy}
    \href{https://docs.scipy.org/doc/scipy/reference/tutorial/general.html}{SciPy Documentation}. \textit{scipy.org}. Retrieved 2019-05-05.
    
    \bibitem{Haar}
    Kasinski A, Schmidt A. \textit{The Architecture of the Face and Eyes DetectionSystem Based on Cascade Classifier.} October 2007, DOI: 10.1007/978-3-540-75175-5\textunderscore16, part of the Advances in Soft Computing book series (AINSC, volume 45)
    
    \bibitem{Viola}
    Viola P, Jones M. \textit{Rapid object detection using a boosted cascade of simple features.} 2001, in: Proceedings of CVPR 1:511–518.
    
    \bibitem{Lienhart}
    Lienhart R, Kuranov A, Pisarevsky V. \textit{Empirical Analysis of Detection Cascades of Boosted Classifiers for Rapid Object Detection.} Technical report, Microprocessor Research Lab, Intel Labs, 2002.
    
    \bibitem{Meynet}
    Meynet J, Popovici V, Thiran J. \textit{Face Detection with Mixtures of Boosted Discriminant Features}. Technical report, EPFL, 2005.
    
    \bibitem{Wang}
    Wang Q, Yang J. \textit{Eye Detection in Facial Images with Unconstrained Background}. Journal of Pattern Recognition Research, 1:55–62, 2006.
    
    \bibitem{OpenViola}
    A. Mohsen Abdul Hossen, R. Abd Alsaheb Ogla, M. Mahmood Ali. \textit{Face Detection by Using OpenCV's Viola-Jones Algorithm based on coding eyes}.November 2017,Iraqi Journal of Science 58(58):735-745.
    
    \bibitem{Fernandez}
    Wilson P., Fernandez J. \textit{Facial feature detection using Haar classifiers.} J. Comput. Small Coll. 21:127–133, 2006.
 
    \bibitem{MMIST}
    Yann LeCun, Corinna Cortes, Christopher J.C. Burges. \textit{THE MNIST DATABASE of handwritten digits}.
    \href{http://yann.lecun.com/exdb/mnist/}{MMIST  database}. mnist.
    
    \bibitem{Reinforcement}
    Richard S. Sutton,Andrew G. Barto. \textit{Reinforcement Learning: An Introduction}. The MIT Press Cambridge, Massachusetts London, England. Second edition, in progress, complete draft. January 1, 2018.
    
    \bibitem{Metrics}
    Mohammed Sunasra. \textit{Performance Metrics for Classification problems in Machine Learning}. https://medium.com/thalus-ai/performance-metrics-for-classification-problems-in-machine-learning-part-i-b085d432082b.
    
    \bibitem{Evaluate}
    Aditya Mishra. \textit{Metrics to Evaluate your Machine Learning Algorithm}. https://towardsdatascience.com/metrics-to-evaluate-your-machine-learning-algorithm-f10ba6e38234.
    
    \bibitem{Review}
    Nurulhuda Ismail, Mas Idayu Md. Sabri. \textit{Review  of  Existing  Algorithms  for  Face  Detection  and  Recognition.} Proceedings of the 8th WSEAS International Conference on  Recent Advances in Computational  Intelligence, Man-Machine Systems and Cybernetics, Puerto De La Cruz, Tenerife, Canary Islands,  Spain, December 14-16, 2009, pp. 30-39.
    
    \bibitem{Survey}
    Hiyam Hatem, Zou Beiji, Raed Majeed. \textit{A Survey of Feature Base Methods for Human Face Detection.} International Journal of Control and Automation Vol.8, No.5 (2015), pp.61-78.
\end{thebibliography}

\end{document}
