\textcolor{green}{\sout{teu cred ca la Requirements ar tb sa imparti in 2 task-urile: unele evidente (functionalitatile aplciatiei: monitorizare si avertizare driver) si altele ascunse (componentele si modelele inteligente (bazate pe ANN-uri)}}

\begin{itemize}
    \item \textbf{Problem statement:} Create an intelligent system that is capable to detect and warn the driver whether him/her is paying attention to the road or not and a drowsiness system that tells if the driver is about to fall asleep.
    \item \textbf{Requirements:} 
        \begin{itemize}
            \item \textbf{Functional:} \begin{enumerate}
                \item create and implement a drowsiness detector to alert the driver to not fall asleep.
                \item Monitorize at any time and check if the driver is paying attention to the road.
                \item Make sure that if the vehicle is stationary not to signal anything.
                \item If the face is not detected make sure to alert the driver.
                \item If the face is detected predict, to check whether the driver is focused, alert if is not paying attention.
            \end{enumerate}
            \item \textbf{Non-functional:} \begin{enumerate}
                \item Create and implement a neural network in order to detect if the driver is focused.
                \item Find a suitable dataset with photos of human faces, that contains a lot of images in order to have diversity and distribute them in labels.
                \item Detect the face and eyes and train the network on those images.
                \item Create and implement an user friendly UI and architecture.
            \end{enumerate}
        \end{itemize}
\end{itemize}